---
title: "ICES Stocks"
subtitle: "Data Sets"
author: "L Kell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
mathjax: TRUE
tags: FLR FLCore introduction
license: Creative Commons Attribution-ShareAlike 4.0 International Public License
#bibliography: ../refs.bib
#knit: (function(inputFile, encoding) {
#  rmarkdown::render(inputFile, encoding = encoding, output_dir = "../outputs/html") })
---


ICES Category 1 stock assessments, analytical age based assessments that have been benchmarked.

```{r, knitr, eval=TRUE, warning=FALSE, cache=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(knitr)
opts_chunk$set(comment   =NA, 
               warning   =FALSE, 
               message   =FALSE, 
               error     =FALSE, 
               echo      =FALSE, 
               eval      =TRUE,
               cache     =TRUE,
               cache.path="../cache/icesStocks/",
               fig.path  ="../figs/icesStocks",
               fig.width =10,
               fig.height=12,
               dev       ="png")

iFig=0
iTab=0
```

```{r}
library(plyr)
library(dplyr)
library(reshape) 
library(ggplot2)

library(FLCore)  
library(FLasher)
library(FLBRP)
library(ggplotFL)
library(FLife)

#library(FLSRTMB)

library(ggpubr)
library(scales) 

library(statcomp)
library(DescTools)
```

  
```{r, eval=!FALSE}
setwd("C:/active/icesdata/Rmd")  
```

```{r, source}
source("../source/jabba-utils.R")
source("../source/spr0Yr.R")
source('../source/nonStationarity.R')
```


```{r, data}
load(         "../data-raw/Updated_stks_n82_R0_updated202408.RData")
info=read.csv("../data-raw/inputs/Database.csv")[,-9]
names(info)[1]=".id"

save(info,file="../data/info.RData")
```

### Meta-data

+ **Attributes** `FLStock` objects with atrributes containing ICES benchmarks and `FishLife` lifehistory parameters
+ **info** `data.frame` with ICES stock codes, expert Working Groups and stock assessment method
+ **spp** `data.frame` with ICES stock codes, common and latin names, and expert Working Groups
+ **`lw`** length weight relationships

```{r, spp}
stock=c("ank.27.78abd", "aru.27.5a14", "aru.27.5b6a", "bss.27.4bc7ad-h", "bss.27.8ab", 
        "cod.21.1", "cod.2127.1f14", "cod.27.1-2", "cod.27.1-2coastN", "cod.27.22-24", 
        "cod.27.24-32", "cod.27.47d20", "cod.27.5a", "cod.27.5b1", "cod.27.6a", 
        "cod.27.7a", "cod.27.7e-k", "had.27.1-2", "had.27.46a20", "had.27.5a", 
        "had.27.5b", "had.27.6b", "had.27.7a", "had.27.7b-k", "her.27.1-24a514a", 
        "her.27.20-24", "her.27.25-2932", "her.27.28", "her.27.3031", "her.27.3a47d", 
        "her.27.5a", "her.27.6a7bc", "her.27.irls", "her.27.nirs", "hke.27.3a46-8abd", 
        "hke.27.8c9a", "hom.27.2a4a5b6a7a-ce-k8", "hom.27.9a", "ldb.27.8c9a", "lin.27.5a", 
        "lin.27.5b", "mac.27.nea", "meg.27.7b-k8abd", "meg.27.8c9a", "mon.27.78abd", 
        "mon.27.8c9a", "pil.27.8abd", "pil.27.8c9a", "ple.27.21-23", "ple.27.420", 
        "ple.27.7a", "ple.27.7d", "pok.27.1-2", "pok.27.3a46", "pok.27.5a", 
        "pok.27.5b", "pra.27.3a4a", "reb.27.1-2", "reg.27.1-2", "reg.27.561214", 
        "san.sa.1r", "san.sa.2r", "san.sa.3r", "san.sa.4", "sol.27.20-24", 
        "sol.27.4", "sol.27.7a", "sol.27.7d", "sol.27.7e", "sol.27.7fg", 
        "sol.27.8ab", "spr.27.22-32", "spr.27.3a4", "tur.27.4", "usk.27.5a14", 
        "whb.27.1-91214", "whg.27.47d", "whg.27.6a", "whg.27.7a", "whg.27.7b-ce-k", 
        "wit.27.3a47d")

name=c("Anglerfish", "Atlantic redfishes", "Atlantic redfishes", "European seabass", 
       "European seabass", "Atlantic cod", "Atlantic cod", "Atlantic cod", 
       "Atlantic cod", "Atlantic cod", "Atlantic cod", "Atlantic cod", 
       "Atlantic cod", "Atlantic cod", "Atlantic cod", "Atlantic cod", 
       "Atlantic cod", "Haddock", "Haddock", "Haddock", 
       "Haddock", "Haddock", "Haddock", "Haddock", "Atlantic herring", 
       "Atlantic herring", "Atlantic herring", "Atlantic herring", 
       "Atlantic herring", "Atlantic herring", "Atlantic herring", 
       "Atlantic herring", "Atlantic herring", "Atlantic herring", 
       "European hake", "European hake", "Atlantic horse mackerel", 
       "Atlantic horse mackerel", "Four-spot megrim", "Ling", 
       "Ling", "Mackerel", "Megrim", "Megrim", "Anglerfish", 
       "Anglerfish", "European pilchard", "European pilchard", 
       "European plaice", "European plaice", "European plaice", 
       "European plaice", "Saithe", "Saithe", "Saithe", 
       "Saithe", "Northern prawn", "Redfish", "Redfish", "Redfish", 
       "Sandeel", "Sandeel", "Sandeel", "Sandeel", "Sole", 
       "Sole", "Sole", "Sole", "Sole", "Sole", 
       "Sole", "Sprat", "Sprat", "Turbot", "Tusk", 
       "Blue whiting", "Whiting", "Whiting", "Whiting", "Whiting", 
       "Witch flounder")

spp=c("Lophius piscatorius", "Sebastes spp.", "Sebastes spp.", "Dicentrarchus labrax", 
      "Dicentrarchus labrax", "Gadus morhua", "Gadus morhua", "Gadus morhua", 
      "Gadus morhua", "Gadus morhua", "Gadus morhua", "Gadus morhua", 
      "Gadus morhua", "Gadus morhua", "Gadus morhua", "Gadus morhua", 
      "Gadus morhua", "Melanogrammus aeglefinus", "Melanogrammus aeglefinus", "Melanogrammus aeglefinus", 
      "Melanogrammus aeglefinus", "Melanogrammus aeglefinus", "Melanogrammus aeglefinus", "Melanogrammus aeglefinus", "Clupea harengus", 
      "Clupea harengus", "Clupea harengus", "Clupea harengus", 
      "Clupea harengus", "Clupea harengus", "Clupea harengus", 
      "Clupea harengus", "Clupea harengus", "Clupea harengus", 
      "Merluccius merluccius", "Merluccius merluccius", "Trachurus trachurus", 
      "Trachurus trachurus", "Lepidorhombus boscii", "Molva molva", 
      "Molva molva", "Scomber scombrus", "Lepidorhombus whiffiagonis", 
      "Lepidorhombus whiffiagonis", "Lophius piscatorius", 
      "Lophius piscatorius", "Sardina pilchardus", "Sardina pilchardus", 
      "Pleuronectes platessa", "Pleuronectes platessa", "Pleuronectes platessa", 
      "Pleuronectes platessa", "Pollachius virens", "Pollachius virens", "Pollachius virens", 
      "Pollachius virens", "Pandalus borealis", "Sebastes spp.", "Sebastes spp.", "Sebastes spp.", 
      "Ammodytes spp.", "Ammodytes spp.", "Ammodytes spp.", "Ammodytes spp.", "Solea solea", 
      "Solea solea", "Solea solea", "Solea solea", "Solea solea", 
      "Solea solea", "Solea solea", "Sprattus sprattus", 
      "Sprattus sprattus", "Scophthalmus maximus", "Brosme brosme", 
      "Micromesistius poutassou", "Merlangius merlangus", "Merlangius merlangus", "Merlangius merlangus", 
      "Merlangius merlangus", "Glyptocephalus cynoglossus")
Genus=c("Lophius", "Sebastes", "Sebastes", "Dicentrarchus", 
        "Dicentrarchus", "Gadus", "Gadus", "Gadus", 
        "Gadus", "Gadus", "Gadus", "Gadus", 
        "Gadus", "Gadus", "Gadus", "Gadus", 
        "Gadus", "Melanogrammus", "Melanogrammus", "Melanogrammus", 
        "Melanogrammus", "Melanogrammus", "Melanogrammus", "Melanogrammus", "Clupea", 
        "Clupea", "Clupea", "Clupea", 
        "Clupea", "Clupea", "Clupea", 
        "Clupea", "Clupea", "Clupea", 
        "Merluccius", "Merluccius", "Trachurus", 
        "Trachurus", "Lepidorhombus", "Molva", 
        "Molva", "Scomber", "Lepidorhombus", 
        "Lepidorhombus", "Lophius", 
        "Lophius", "Sardina", "Sardina", 
        "Pleuronectes", "Pleuronectes", "Pleuronectes", 
        "Pleuronectes", "Pollachius", "Pollachius", "Pollachius", 
        "Pollachius", "Pandalus", "Sebastes", "Sebastes", "Sebastes", 
        "Ammodytes", "Ammodytes", "Ammodytes", "Ammodytes", "Solea", 
        "Solea", "Solea", "Solea", "Solea", 
        "Solea", "Solea", "Sprattus", 
        "Sprattus", "Scophthalmus", "Brosme", 
        "Micromesistius", "Merlangius", "Merlangius", "Merlangius", 
        "Merlangius", "Glyptocephalus")
spp=data.frame(stock=stock,name=name,spp=spp,Genus=Genus,Species=unlist(strsplit(spp," "))[seq(2,length(spp)*2,2)])

save(spp,file="../data/spp.RData")
```


```{r}
lw=data.frame(
    name=c(
      "Anglerfish","Atlantic redfishes","European seabass","Atlantic cod",
      "Haddock", "Atlantic herring", "European hake", "Atlantic horse mackerel", 
      "Four-spot megrim","Ling","Mackerel","Megrim","European pilchard",
      "European plaice","Saithe","Northern prawn","Redfish","Sandeel", 
      "Sole","Sprat","Turbot","Tusk", "Blue whiting","Whiting","Witch flounder"),
    latin=c(
      "Lophius piscatorius","Sebastes spp.","Dicentrarchus labrax","Gadus morhua", 
      "Melanogrammus aeglefinus","Clupea harengus","Merluccius merluccius","Trachurus trachurus", 
      "Lepidorhombus boscii","Molva molva","Scomber scombrus","Lepidorhombus whiffiagonis", 
      "Sardina pilchardus","Pleuronectes platessa","Pollachius virens","Pandalus borealis", 
      "Sebastes marinus","Ammodytes spp.","Solea solea","Sprattus sprattus", 
      "Scophthalmus maximus","Brosme brosme","Micromesistius poutassou","Merlangius merlangus", 
      "Glyptocephalus cynoglossus"),
    a=c(0.0045,0.0050,0.0075,0.0061,0.0053,0.0032,0.0048,0.0059,0.0046,0.0038,0.0057,
          0.0046,0.0059,0.0038,0.0062,0.0015,0.0050,0.0055,0.0042,0.0050,0.0071,0.0060,0.0052,0.0054,0.0043),
    b=c(3.25,3.10,3.10,3.04,3.15,3.24,3.12,3.08,3.20,3.25,3.10,3.20,3.02,3.20,3.15,2.98,
        3.10,3.05,3.12,3.10,3.15,3.18,3.10,3.15,3.20),
    Source=c(
      "FishBase","FishBase","ResearchGate","FishBase","ScienceDirect","ResearchGate","ResearchGate",
      "ResearchGate","FishBase","FishBase","FishBase","FishBase","ResearchGate","ScienceDirect",
      "FishBase","NAFO","FishBase","FishBase", "FishBase","FishBase","FishBase","FishBase","FishBase","FishBase","FishBase"))

save(lw,file="../data/lw.RData")
```

### Summaries

#### Catch data


```{r, catch-ices}
ctc    =ldply(ICESStocks, function(x) as.data.frame(catch(x),drop=TRUE))

stdz<-function(x){
  x=x-mean(x,na.rm=TRUE)
  x/var(x,na.rm=TRUE)^0.5}

ggplot(ddply(ctc,.(.id), transform, catch=stdz(data)))+
  geom_line(aes(year,catch),col="blue")+
  geom_text(aes(x=1950, y=2, label=.id), 
            hjust=0, vjust=1, size=3, color="black") +
  facet_wrap(~.id,scale="free_y",ncol=4)+
  theme_minimal(16)+
  theme(legend.position="none", 
        strip.text = element_blank(),
        axis.text  = element_blank(),
        axis.ticks = element_blank())+  
  labs(title="Catch",
       x    ="Year",
       y    ="")
```

**Figure `r iFig=iFig+1; iFig`** ICES catches, these are those estimated by the model if
a statistical catch-at-age method is used.


```{r, catch-1903}
ctc1903=read.csv("../data-raw/inputs/NorthSea_stocks.csv")[,1:14]
ctc1903$Discards[is.na(ctc1903$Discards)]=0

ggplot(ddply(ctc1903,.(FishStock), transform, catch=stdz(Catches)))+
  geom_line(aes(Year,catch),col="blue")+
  geom_text(aes(x=1900, y=2, label=FishStock), 
            hjust=0, vjust=1, size=3, color="black") +
  geom_vline(aes(xintercept=1939),colour="red")+
  facet_wrap(~FishStock,scale="free_y",ncol=4)+
  theme_minimal(10)+
  theme(legend.position="none", 
        strip.text   = element_blank(),
        axis.text.y  = element_blank(),
        axis.ticks.y = element_blank())+  
  labs(title="Catch",
       x    ="Year",
       y    ="")
```

**Figure `r iFig=iFig+1; iFig`** Historical catches, from the ICES historical database
rather than from the stock assessment data sets.

```{r, catch-summary}
ctc2=rbind(cbind(What="ICES",ctc),
           cbind(What="1903",transmute(ctc1903,.id=FishStock,year=Year,
                                       data=Catches)))

ggplot(ddply(ctc2,.(.id), transform, catch=stdz(data)))+
  geom_line(aes(year,catch,col=What))+
  geom_text(aes(x=1900, y=2, label=.id), 
            hjust=0, vjust=1, size=3, color="black") +
  geom_vline(aes(xintercept=1939),colour="red")+
  scale_color_manual(values=c("cyan","blue"))+
  facet_wrap(~.id,scale="free_y",ncol=4)+
  theme_minimal(10)+
  theme(legend.position="none", 
        strip.text   = element_blank(),
        axis.text.y  = element_blank(),
        axis.ticks.y = element_blank())+  
  labs(title="Catch",
       x    ="Year",
       y    ="")
```

**Figure `r iFig=iFig+1; iFig`** Comparison of ICES assessment and historical catches.

```{r, catch-3}
maxyear=ddply(ctc,.(.id), with, data.frame(maxyear=max(year)))
ctc2   =merge(ctc2,maxyear,by=".id",all.x=TRUE)
ctc2   =subset(ctc2,is.na(maxyear)|year<=maxyear)

ctc2$What=factor(ctc2$What,levels=c("ICES","1903"))
ctc3=ctc2[do.call("order",ctc2[,c(".id","year","What")]),]
ctc3=ctc3[!duplicated(ctc3[,c(".id","year")]),]
  
ggplot(ddply(ctc3,.(.id), transform, catch=stdz(data)))+
  geom_line(aes(year,catch,col=What))+
  geom_text(aes(x=1900, y=2, label=.id), 
            hjust=0, vjust=1, size=3, color="black") +
  geom_vline(aes(xintercept=1939),colour="red")+
  scale_color_manual(values=c("blue","cyan"))+
  facet_wrap(~.id,scale="free_y",ncol=4)+
  theme_minimal(10)+
  theme(legend.position="none", 
        strip.text   = element_blank(),
        axis.text.y  = element_blank(),
        axis.ticks.y = element_blank())+  
  labs(title="Catch",
       x    ="Year",
       y    ="")
```

**Figure `r iFig=iFig+1; iFig`** Updated catch time series.


Stocks with missing catch-at-age data

```{r, catch}
ctcNA=subset(ldply(ICESStocks, function(x) any(is.na(catch.n(x)))),V1)[,1]

ctcNA
```

### Indices

```{r}
idx=read.csv("../data-raw/inputs/Stock_indices.csv")[,1:5]
names(idx)=c(".id","survey","year","data","ages")
idx$data=an(idx$data)

smry =ddply(idx,  .(.id,survey), with, data.frame(n=length(data)))
smry2=ddply(smry, .(.id), with, data.frame(n=length(survey)))

save(idx,smry,smry2,file="../data/surveys.RData")

ggplot(ddply(idx,.(survey,.id), transform, index=stdz(data)))+
  geom_line(aes(year,index,col=survey))+
  geom_text(aes(x=1950, y=2, label=.id), 
            hjust=0, vjust=1, size=3, color="black") +
  facet_wrap(~.id,scale="free_y",ncol=4)+
  theme_minimal(16)+
  theme(legend.position="none", 
        strip.text = element_blank(),
        axis.text  = element_blank(),
        axis.ticks = element_blank())+  
  labs(title="Indices",
       x    ="Year",
       y    ="")
```


**Figure `r iFig=iFig+1; iFig`** Indices of abundance.

