---
title: "ICES Stocks"
subtitle: "Derived Quantities"
author: "L Kell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
mathjax: TRUE
tags: FLR FLCore introduction
license: Creative Commons Attribution-ShareAlike 4.0 International Public License
#bibliography: ../refs.bib
#knit: (function(inputFile, encoding) {
#  rmarkdown::render(inputFile, encoding = encoding, output_dir = "../outputs/html") })
---


ICES Category 1 stock assessments, analytical age based assessments that have been benchmarked.

```{r, knitr, eval=TRUE, warning=FALSE, cache=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(knitr)
opts_chunk$set(comment   =NA, 
               warning   =FALSE, 
               message   =FALSE, 
               error     =FALSE, 
               echo      =FALSE, 
               eval      =TRUE,
               cache     =TRUE,
               cache.path="../cache/derived/",
               fig.path  ="../figs/derived",
               fig.width =10,
               fig.height=12,
               dev       ="png")

iFig=0
iTab=0
```

```{r, eval=FALSE}
library(remotes)

remotes::install_github("flr/FLSRTMB", ref="Depensation")
```

```{r}
library(plyr)
library(dplyr)
library(reshape) 
library(ggplot2)

library(FLCore)  
library(FLasher)
library(FLBRP)

library(ggplotFL)
library(FLife)

library(FLCandy)

library(FLSRTMB)

library(ggpubr)
library(scales) 

library(statcomp)
library(DescTools)
```
  

```{r, source}
source("../source/jabba-utils.R")
source("../source/spr0Yr.R")
source('../source/nonStationarity.R')
```


```{r, data}
load("../data-raw/Updated_stks_n82_R0_updated202408.RData")
load("../data/info.RData")
load("../data/spp.RData")
```

### Summaries

Biological assumptions and variation by age, and year 

```{r}
vary=ldply(ICESStocks, function(x) data.frame(
                    m.age  =!all(m(x)[1,]==apply(m(x),2,mean)),
                    m.yr   =!all(apply(       m(x),2,mean)==mean(m(       x)[,1])),
                    mass.yr=!all(apply(stock.wt(x),2,mean)==mean(stock.wt(x)[,1])),
                    mat.yr =!all(apply(     mat(x),2,mean)==mean(mat(     x)[,1]))))
vary=transform(vary,m.by.age  =ifelse(m.age,"M varys by Age",     "M constant by Age"),
                    m.by.yr   =ifelse(m.yr, "M varys by Year",    "M constant by Year"),
                    mat.by.yr =ifelse(mat.yr, "Mat varys by Year",  "Mat constant by Year"),
                    mass.by.yr=ifelse(mass.yr, "Mass varys by Year", "Mass constant by Year"))

save(vary,file="../data/vary.RData")
```


### Equilibrium analysis

Take the average values=at-age across all years for each stock, and fit a 
Beverton and Holt stock recruitment relationship using steepness from `FishBase'

```{r, eq}
eq=mlply(names(ICESStocks), function(.id) {
  rtn=try(eqlFn(ICESStocks[[.id]]))
  
  if ("try-error"%in%is(rtn)) return(NULL)
  
  return(rtn)})

names(eq)=names(ICESStocks)
save(eq,file="../data/eq.RData")
```

### Process error

Process error represents the inherent variability in population dynamics that is 
not accounted for by the model structure. It differs from sampling error, which 
arises from the data collection process.  However, it can be difficult to consistently 
identify the correct process error structure. Incorrectly attributing process error to 
the wrong model component (e.g. natural mortality vs selectivity) can lead to 
biased estimates and management advice.


```{r, pe}
pe=mdply(names(ICESStocks), function(.id) {  
    
  Fn<-function(stk,eq,stock=FLCore:::ssb){
          (stock(stk)%-%
          window(stock(stk)[,-1],end=dims(stk)$maxyear+1)-
          catch(stk)%+%sp(stk,eq,stock))%/%stock(stk)}
  
     rtn=try(Fn(ICESStocks[[.id]],eq[[.id]])) 
     if ("try-error"%in%is(rtn)) return(NULL) else 
       cbind(.id=.id,as.data.frame(rtn,drop=TRUE))})[,-1]

ggplot(pe)+
  geom_line(aes(year,data),col="blue")+
  geom_text(aes(x=1950, y=V1, label=.id), data=ddply(pe,.(.id), with,    
                                                     median(data,na.rm=TRUE)),
              hjust=0, vjust=1, size=3, color="black") +
  facet_wrap(~.id,scale="free_y",ncol=4,)+
  theme_minimal(10)+
  theme(legend.position="none", 
        strip.text   = element_blank(),
        axis.text.y  = element_blank(),
        axis.ticks.y = element_blank())+  
  labs(title="Process Error",
       x    ="Year",
       y    ="")

save(pe,file="../data/pe.RData")
```
**Figure `r iFig=iFig+1; iFig`** Process error,


#### $SPR_{0}$ time series


```{r, spr0}
spr0Par=llply(ICESStocks, function(x) {
  res=spr0Yr(x)
  FLPar(array(c(res),dim=c(1,dim(res)[2]),dimnames=list(params="spr0",iter=seq(dim(res)[2]))))})

spr0=ldply(ICESStocks, function(x) as.data.frame(spr0Yr(x),drop=T))  

ggplot(spr0)+
  geom_line(aes(year,data),col="blue")+
  geom_text(aes(x=1950, y=V1, label=.id), data=ddply(spr0,.(.id), with,    
                                                     median(data,na.rm=TRUE)),
              hjust=0, vjust=1, size=3, color="black") +
  facet_wrap(~.id,scale="free_y",ncol=4,)+
  theme_minimal(10)+
  theme(legend.position="none", 
        strip.text   = element_blank(),
        axis.text.y  = element_blank(),
        axis.ticks.y = element_blank())+  
  labs(title="SPR0",
       x    ="Year",
       y    ="")

save(spr0,spr0Par,file="../data/spr0.RData")
```

#### SRR

```{r, sr, eval=FALSE}
sr=FLSRs(lapply(ICESStocks, function(x) { 
  
  srr = as.FLSR(x,model=bevholtSV)
  
  res=ftmb(srr,s.est=T,s=x@fishlife["s"],s.logitsd=x@fishlife["sd.logit.s"],spr0=spr0Yr(x))
  res}))
names(sr)=names(ICESStocks)

srPar=FLPars(mlply(names(sr), function(x){

  res=spr0Yr(ICESStocks[[x]])
  res=FLPar(array(c(res),dim=c(1,dim(res)[2]),dimnames=list(params="spr0",iter=seq(dim(res)[2]))))
  
  rbind(params(sr[[x]]),res)
  }))
names(srPar)=names(ICESStocks)

for (i in seq(length(srPar))) 
  srPar[[i]]=try(FLCore:::sv(srPar[[i]],model="bevholt"))

save(sr,srPar,file="../data/sr.RData")
```

```{r, virgin}
virgin=mdply(names(ICESStocks), function(.id) {
  cbind(.id=.id,model.frame(srPar[[.id]]),year=unlist(dimnames(spr0Yr(ICESStocks[[.id]]))$year))})[,-c(1,6)]
virgin$year=as.numeric(virgin$year)

ggplot(virgin)+
  geom_line(aes(year,v),col="blue")+
  geom_text(aes(x=1950, y=V1, label=.id), data=ddply(virgin,.(.id), with,    
                                                     median(v,na.rm=TRUE)),
             hjust=0, vjust=1, size=3, color="black") +
  facet_wrap(~.id,scale="free_y",ncol=4)+
  theme_minimal(10)+
  theme(legend.position="none", 
        strip.text   = element_blank(),
        axis.text.y  = element_blank(),
        axis.ticks.y = element_blank())+  
  labs(title="Virgin Biomass",
       x    ="Year",
       y    ="")

save(virgin,file="../data/virgin.RData")
```

\newpage

### Reference Points

```{r, msy-data}
rf=mdply(names(ICESStocks), function(.id){

    rtn=cbind(.id=.id,nonStationarity(ICESStocks[[.id]],sr=sr[[.id]]))
    if ("try-error"%in%is(rtn)) return(NULL)
    rtn})
rf=cast(rf,.id+year+refpt~quant,value="data")
```


```{r, msy-plot}
ggplot(subset(rf,refpt=="msy"))+
  geom_line(aes(year,ssb),col="blue")+
  geom_text(aes(x=1950, y=V1, label=.id), data=ddply(rf,.(.id), with,    
                                                     median(ssb,na.rm=TRUE)),
             hjust=0, vjust=1, size=3, color="black") +
  facet_wrap(~.id,scale="free_y",ncol=4)+
  theme_minimal(10)+
  theme(legend.position="none", 
        strip.text   = element_blank(),
        axis.text.y  = element_blank(),
        axis.ticks.y = element_blank())+  
  labs(title="BMSY",
       x    ="Year",
       y    ="")

save(rf,file="../data/nonStatRefs.RData")
```


```{r}

```