---
title: "ICES Stocks"
subtitle: "Production Functions"
author: 
  L.T. Kell
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  pdf_document:
    toc: true
    number_sections: true
vignette: >
  %\VignetteIndexEntry{Estimating $F_{MNY}$ for Atlantic Mackerel}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

# Introduction


Data are in the `icesdata` package

```{r, icesdata, eval=FALSE}
library(icesdata)

data(icesdata)
```

```{r, knitr, eval=TRUE, echo=FALSE, warning=FALSE, cache=FALSE}
library(knitr)

opts_chunk$set(comment   =NA, 
               warning   =FALSE, 
               message   =FALSE, 
               error     =FALSE, 
               echo      =FALSE, 
               cache     =TRUE, 
               cache.path="../cache/pFn/",
               fig.path  ="../figs/pFn",
               fig.width =8,
               fig.height=6,
               dev       ="png")

iFig=0
iTab=0
```

```{r, pkgs}
library(FLCore)
library(FLBRP)
library(FLasher)
library(FLife)

require(ggplot2)
require(ggpubr)
require(ggplotFL)

library(GGally)
library(ggpubr)

library(plyr)
library(dplyr)
library(reshape)
library(kableExtra)

library(foreach)
library(doParallel)

theme_set(theme_bw(16))
```

```{r, eval=FALSE}
setwd("C:/flrpapers/jabba_eval/Rmd")
```

```{r, source}
source("../source/jabba-utils.R", echo=TRUE)
```

```{r, ICES-stocks}
load("C:/active/icesdata/data/icesdata.RData")
load("C:/active/icesdata/data/spp.RData")
```


```{r, tseries, echo=TRUE}
# Use all cores except four
ncore=detectCores() - 4  
registerDoParallel(ncore)

stks=names(icesdata)

bevHoltSV=foreach(id=stks, .combine   =list, 
                         .maxcombine=length(stks), 
                         .packages  =c("FLCore","FLBRP","plyr","dplyr"),
                         .export    =c("icesdata","eqlFn")) %dopar% {
  try(eqlFn(icesdata[[id]], model="bevholtSV"))}
bevHoltSV=FLBRPs(bevHoltSV)
names(bevHoltSV)=stks
save(bevHoltSV,file="../data/bevHoltSV.RData")

bevHolt=foreach(id=stks, .combine   =list, 
                         .maxcombine=length(stks), 
                         .packages  =c("FLCore","FLBRP","plyr","dplyr"),
                         .export    =c("icesdata","eqlFn")) %dopar% {
  try(eqlFn(icesdata[[id]], model="bevholt"))}
bevHolt=FLBRPs(bevHolt)
names(bevHolt)=stks
save(bevHolt,file="../data/bevHolt.RData")

ricker=foreach(id=stks, .combine   =list, 
                         .maxcombine=length(stks), 
                         .packages  =c("FLCore","FLBRP","plyr","dplyr"),
                         .export    =c("icesdata","eqlFn")) %dopar% {
  try(eqlFn(icesdata[[id]], model="ricker"))}
ricker=FLBRPs(ricker)
names(ricker)=stks
save(ricker,file="../data/ricker.RData")

stopCluster(ncore)
```

```{r, catch-ices, fig.height=10, fig.width=8}
ggplot(ddply(tseries,.(.id), transform, catch=stdz(catch)))+ 
  geom_line(aes(year,catch),col="blue")+
  geom_text(aes(x=1950, y=2, label=.id), 
            hjust=0, vjust=1, size=3, color="black") +
  facet_wrap(~.id,scale="free_y",ncol=4)+
  theme_minimal(16)+
  theme(legend.position="none", 
        strip.text = element_blank(),
        axis.text  = element_blank(),
        axis.ticks = element_blank())+  
  labs(title="Catch: Cat 1 stocks", 
       x    ="Year",
       y    ="")
```
**Figure `r iFig=iFig+1; iFig`** Time series of standardised catches from Cat 1 stocks


```{r, ssb-ices, fig.height=10, fig.width=8}
ggplot(ddply(tseries,.(.id), transform, ssb=stdz(ssb)))+
  geom_line(aes(year,ssb),col="blue")+
  geom_text(aes(x=1950, y=2, label=.id), 
            hjust=0, vjust=1, size=3, color="black")+
  facet_wrap(~.id,scale="free_y",ncol=4)+
  theme_minimal(16)+
  theme(legend.position="none", 
        strip.text = element_blank(),
        axis.text  = element_blank(),
        axis.ticks = element_blank())+  
  labs(title="SSB: Cat 1 stocks",
       x    ="Year",
       y    ="")
```
**Figure `r iFig=iFig+1; iFig`** Time series of standardised SSB from Cat 1 stocks


```{r, fig.height=8, fig.width=8}
my_smooth <- function(data,mapping,...){
  ggplot(data=data,mapping=mapping)+
  geom_point(...,size=.5)+
  geom_smooth(...,method="lm",se=FALSE)}

my_density <- function(data,mapping,...){
  ggplot(data=data,mapping=mapping)+
  geom_density(...,lwd=1)}
  
ggpairs(model.frame(priors[,c("r","fmsy","shape")]),
  lower = list(continuous = wrap(my_smooth)),
  diag=list(continuous=wrap(my_density,alpha=0.2)),
  title = "")+
  theme(legend.position ="bottom",
  panel.grid.major =element_blank(),
  axis.ticks       =element_blank(),
  axis.text.x      =element_text(angle=-45),
  axis.text.y      =element_blank(),
  panel.border     =element_rect(linetype = 1, colour="black", fill=NA))+
  theme_bw()
```

**Figure `r iFig=iFig+1; iFig`** Relationships between priors


```{r, run-perfect, eval=FALSE}
# Use all cores except four
ncore=detectCores() - 4  
registerDoParallel(ncore)

stks=names(icesdata)[names(icesdata)%in%priors$.id]

perfect=foreach(id=stks, .combine   =list, 
                         .maxcombine=length(stks), 
                         .packages  =c("dplyr","JABBA"),
                         .export    =c("jabFn","tseries","priors")) %dopar% {

    pr=subset(priors,.id==id)           
    ts=merge(subset(tseries, .id==id),pr, by=".id")
                                
    if (is.na(pr["r"])) pr["r"]=0.3
                                
    catch=ts[, c("year", "catch")]
    index=transmute(ts, year=year, index=eb)
  
  try(jabFn(catch,pr,index,q_bound=c(0.25,2),pr.sd=c(0.1,0.1),
            assessment=id,scenario="perfect"))}
names(perfect)=stks

stopCluster(ncore)

save(perfect,file="../data/results/perfect.RData")

perfectFits=cbind(scen="Perfect",tsFn(perfect))
perfectFits$year=as.numeric(as.character(perfectFits$year))
save(perfectFits,file="../data/results/perfectFits.RData")

perfectKobe=cbind(scen="Perfect",kbFn(perfect))
save(perfectKobe,file="../data/results/perfectKobe.RData")
```

```{r, plot-perfect, eval=FALSE}
id=names(perfect)[1]
jbplot_summary(list("Perfect"=perfect[[id]][[2]]))
```



```{r, plot-all, fig.height=10, fig.width=8}
load("../data/results/perfectFits.RData")

ggplot(perfectFits)+ 
  geom_line(aes(year,BBmsy),col="red")+ 
  geom_line(aes(year,ssb/bmsy), col="orange",data=merge(tseries,priors))+
  geom_hline(aes(yintercept=1.0))+
  geom_text(aes(x=1980, y=2, label=.id), 
            hjust=0, vjust=1, size=3, color="black")+
  facet_wrap(~.id,scale="free",ncol=6)+ 
  labs(title="B/BMSY",
       x    ="",
       y    ="")+ 
  theme_minimal()+
  theme(legend.position="bottom",
        strip.text = element_blank(),
        axis.text  = element_blank(),
        axis.ticks = element_blank()) 
```

**Figure `r iFig=iFig+1; iFig`** Estimates of $B/B_{MSY}$ from JABBA (red) compared with Cat 1 assessments (orange).


```{r, plot-kobe, fig.height=10, fig.width=8}
load("../data/results/perfectKobe.RData")

ggplot(perfectKobe)+
  geom_histogram(aes(stock))+
  geom_vline(aes(xintercept=ssb.maxyr/bmsy),data=priors,col="red")+
  geom_text(aes(x=1,y=500,label=.id), hjust=0, vjust=1, size=3, color="blue")+
  facet_wrap(~.id,scale="free_y",ncol=6)+ 
  labs(title="B/BMSY",
       x    ="",
       y    ="")+ 
  theme_minimal()+
  theme(legend.position="bottom",
        strip.text = element_blank(),
        axis.text  = element_blank(),
        axis.ticks = element_blank())+
  scale_x_continuous(limits=c(0,2.5))
```

**Figure `r iFig=iFig+1; iFig`** Estimates of $B/B_{MSY}$ in the final year from JABBA (red) compared with Cat 1 assessments (red).


```{r, plot-kobe-harvest, fig.height=10, fig.width=8}
load("../data/results/perfectKobe.RData")

ggplot(perfectKobe)+
  geom_histogram(aes(harvest))+
  geom_vline(aes(xintercept=f.maxyr/fmsy),data=priors,col="red")+
  geom_text(aes(x=1,y=500,label=.id), hjust=0, vjust=1, size=3, color="blue")+
  facet_wrap(~.id,scale="free",ncol=6)+ 
  labs(title="F/FMSY",
       x    ="",
       y    ="")+ 
  theme_minimal()+
  theme(legend.position="bottom",
        strip.text = element_blank(),
        axis.text  = element_blank(),
        axis.ticks = element_blank()) 
```

**Figure `r iFig=iFig+1; iFig`** Estimates of $F/F_{MSY}$ in the final year from JABBA (red) compared with Cat 1 assessments (red).